---
title: "P8105_hw3_gl2677"
author: "Gaotong LIU"
date: "10/5/2019"
output: github_document
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(kableExtra)
library(p8105.datasets)


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```

## Problem 1
There are 1,384,617 observations of 131,209 unique users and 15 vaiables in the `instacart` dataset, where each row is a product from an order. The key variables are the ID and name of product, aisle, department, together with the hour of the day and the day of the week on which the order was placed.

Here is an illstrative example of observations. The first 8 rows are from the same order with ID numerber `1` for the user with ID numerber `112108`. This is the `4`th order this user has placed, and it has been `9` days since the last order. In this order, there are total 8 products. The `1`st product added to cart was `Bulgarian Yogurt` with ID number `49302` from `yogurt` aisle with ID number`120`, `dairy eggs` department with ID number `16`. This producthas been ordered by this user in the past and it was placed on `10` o'clock, `Thursday` this time. 

```{r number of aisle, eval=FALSE}
data("instacart")
instacart %>% 
  group_by(aisle) %>% 
  summarize(n = n()) %>% 
  arrange(desc(n))
```

* There are totally 134 aisles in the dataset. `fresh vegetables` are the most items ordered from, and the majority of items are ordered from `fresh vegetables` and `fresh fruits`.

```{r plot}
instacart %>% 
  group_by(aisle) %>% 
  summarize(n = n()) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n, fill = aisle)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "The number of items ordered in each aisle",
       x = "Aisle",
       y = "The number of items") +
  theme(legend.position = "none")
```

* From the plot, it shows most of items ordered in `fresh vegetables`, `fresh fruits`, and `packaged vegetables fruits`, and totally around 375,000 items are oredered in these three aisles. People mainly use The Instacart Online Grocery Shopping to buy vegetables and fruits in 2017.

```{r table1}
instacart %>% 
  filter(aisle == c("baking ingredients","dog food care", "packaged vegetables fruits")) %>%
  group_by(product_name, aisle) %>% 
  summarize(n = n()) %>% 
  group_by(aisle) %>% 
  filter(min_rank(desc(n)) <= 3) %>%
  arrange(aisle) %>% 
  ungroup() %>%
  select(-aisle) %>% 
  pivot_wider(names_from = product_name,
              values_from = n) %>% 
  knitr::kable(caption = "The number of times each item is ordered against the three most popular items in each aisle.") %>% 
  add_header_above(c("baking ingredients" = 3,"dog food care" = 3, "packaged vegetables fruits" = 3))
```

* `packaged vegetables fruits` is the most frequently ordered aisles, and `dog food care` is the least frequently ordered ones. Organic ingredients are often picked in these three aisles.

```{r table2}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>% 
  summarize(mean_h = round(mean(order_hour_of_day))) %>% 
  mutate(order_dow = recode(order_dow, `0` = "Sunday",
                            `1` = "Monday",
                            `2` = "Tuesday",
                            `3` = "Wednesday",
                            `4` = "Thursday",
                            `5` = "Friday",
                            `6` = "Saturday"),
         mean_h = str_c(mean_h, "00", sep = ":")) %>%
  pivot_wider(names_from = order_dow,
              values_from = mean_h) %>% 
  knitr::kable(caption = "The mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered") 
```

* Most of `Coffee Ice Cream` is ordered between two o'clock and three o'clock in the afternoon during the week. `Pink Lady Apples` is ordered with unfixed time, which is between eleven o'clock in the morning and two o'clock in the afternoon.

## Problem 2
```{r clean data}
data("brfss_smart2010")
overall_health = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health" & response !="Very good") %>% 
  mutate(response = forcats::fct_relevel(response, c("Poor", "Fair", "Good",  "Excellent")),
         state = locationabbr,
         county = locationdesc) 
```

```{r states}
states_2002 = overall_health %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  filter(n >= 7) %>% 
  arrange(desc(n)) 

states_2010 = overall_health %>% 
  filter(year == 2010) %>% 
  group_by(state) %>% 
  summarize(n = n()) %>% 
  filter(n >= 7) %>% 
  arrange(desc(n)) 
```
* In 2002, total `36` states ([most locations first]``r pull(states_2002, state)``) were observed at 7 or more locations. The range of the number of locations in the state is from `r max(pull(states_2002, n))` to `r min(pull(states_2002, n))`.
* In 2010, total `45` states ([most locations first]``r pull(states_2010, state)``) were observed at 7 or more locations. The range of the number of locations in the state is from `r max(pull(states_2010, n))` to `r min(pull(states_2010, n))`.
* The maximum number of locations in the state increase in 2010 comparing to 2002. 

```{r excellent}
overall_health %>% 
  filter(response == "Excellent") %>% 
  group_by(state,year) %>% 
  summarize(mean_data_value = mean(data_value)) %>% 
  arrange(mean_data_value) %>% 
  ggplot(aes(x = year, y = mean_data_value, color = state)) +
  geom_line() +
  labs(title = "Average data value over time within a state",
       subtitle = "limited to Excellent responses",
       x = "Year",
       y = "Mean data-value(%)")
```

* Most of states share the similar pattern of mean `data_value`, and it fluctuates between 15% and 30%. However, mean`data_value` of state `WV` in 2005 and 2009, and state `IN` in 2007 is obviously lower than 15%.

```{r two-panel}
data_value_2006 = overall_health %>%
  filter(year == 2006) %>% 
  filter(state == "NY") %>% 
  group_by(response) %>% 
  ggplot(aes(x = county, y = data_value, 
             color = response, group = response)) +
  geom_point() +
  geom_line() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "none") +
  labs(title = "Distribution of data-value for responses among locations in NY state in 2006",
       x = "",
       y = "Data-value(%)")

data_value_2010 = overall_health %>%
  filter(year == 2010) %>% 
  filter(state == "NY") %>% 
  group_by(response) %>% 
  ggplot(aes(x = county, y = data_value, 
             color = response, group = response)) +
  geom_point() +
  geom_line() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  labs(title = "Distribution of data-value for responses among locations in NY state in 2010",
       x = "",
       y = "Data-value(%)")
data_value_2006 / data_value_2010
```

* Most of counties in NY state share the similar pattern that `Poor` response associates with lowest `data_value`, `Good` response associates with highest `data_value`. `Excellent` response has higher value than `Fair` response, but lower than` Good` response both in 2006 and 2010. 
* Only New York County has different pattern both in 2006 and 2010. For New York County, `Excellent` response has the highest value ,and `Good`, `Fair` and `Poor` responses are descending in order.

## Problem 3
```{r wrangle data, message=FALSE}
activity = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(activity_1:activity_1440, 
               names_to = "minute",
               names_prefix = "activity_",
               values_to = "minute_activity") %>% 
  mutate(minute = as.numeric(minute),
         week = as.factor(week),
         hour = minute %/% 60,
         weekday = factor(day %in% c("Sunday", "Saturday"), 
                          levels = c(TRUE, FALSE), 
                          labels = c("weekend", "weekday")),
         day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday","Sunday")))   
```
* There are `r nrow(activity)` observations and `r ncol(activity)` variables in `acticity` dataset. The variables are ``r colnames(activity)``. For each observation, the dataset shows the activity counts for the exact minute of a 24-hour day starting at midnight on exact day and exact week.

```{r day_activity table}
activity %>% 
  group_by(week, day) %>% 
  summarize(day_activity = sum(minute_activity)) %>% 
  pivot_wider(names_from = day,
              values_from = day_activity) %>% 
  knitr::kable(digits = 0,
               caption ="The total activity over the day for five weeks") 

 activity %>% 
  group_by(week, day) %>% 
  summarize(day_activity = sum(minute_activity)) %>% 
  ggplot(aes(x = day, y = day_activity, color = week, group = week)) +
  geom_point() +
  geom_line() +
  labs(title = "The total activity over the day for five weeks",
       x = "",
       y = "Day activity counts")
```

* The accelerometer acticity counts for all five weeks increase on Wednesday comparing to Tuesday. and they all are numerically similar between Tuesday and Thursay. The activity counts on Saturday drop to the same value in week 4 and week 5.

```{r minute_activity plot}
activity %>% 
  group_by(hour, day) %>% 
  summarize(mean_hour_activity = mean(minute_activity)) %>% 
  ggplot(aes(x = hour, y = mean_hour_activity,
             color = day, group = day)) +
  geom_point() +
  geom_line() +
  labs(title = " 24-hour activity time courses for each day ",
       x = "Hour in the day",
       y = "Mean hour activity counts")
```

* The `mean_hour_activity` except on Sunday has the similar pattern that the counts increase starting from around 5:00 and reach the peak around 20:00, then decrease until 23:00. This means the most intense physical activity occurs around eight o'clock in the evening, and he might get up around five o'clock in the morning and go to bed around eleven o'clock in the evening.
* The `mean_hour_activity` counts on Sunday increase starting from around 5:00 and reach the peak around 11:00, then decrease until 23:00. This means the most intense physical activity occurs around 11 o'clock in the morning, and he might get up around five o'clock in the morning and go to bed around eleven o'clock in the evening.

